AFQ-Insight

### 2019-02-11 reports/manuscript/sections/introduction.tex
These are better references for predictive model "culture".

Diffusion-weighted Magnetic Resonance Imaging (dMRI) provides a unique view into
the physical properties of the connections that comprise the brain white matter
\cite{wandell2016clarifying}. Among the benefits of the method are the fact that
it is non-invasive, and feasible to conduct in clinical scanners, and in wide
variety of populations: healthy adults, but also clinical populations, as well
as young or elderly subjects. While the measurements are usually conducted with
voxels at the millimeter scale, within the diffusion time usually used, water
molecules within each voxel can probe the structure of white matter at the
micrometer scale, providing important information about the physical structure
of the white matter, such as the distribution of fiber orientations within each
voxel \cite{Frank2002-iz, Frank2001-xf}. With more sophisticated measurements
that include several diffusion-weightings, more information can be extracted
about the composition of different tissue type in the voxel
\cite{Jeurissen2014-ab, Kaden2016-ho}. In addition, phenomenological
descriptions of the diffusion properties, such as its mean diffusivity or
fractional anisotropy can be derived from dMRI measurements
\cite{Pierpaoli1996-vj}. Some of these descriptions are ambiguous in terms of
their interpretation, as variance in these properties can relate to several
different underlying physical causes. For example, fractional anisotropy can
decrease both because of demyelination, but also because of decreased density,
or because of partial volume with other tissue types or between crossing fiber
populations \cite{Beaulieu1996-fn, Beaulieu2002-tl}. Even so, these descriptions
have proven useful in characterizing individual differences or differences
between populations \cite{}.

To relate the diffusion in each voxel to the underlying macro-structure of
long-range connections between different brain regions, methods for
computational tract-tracing from diffusion MRI, or tractography combine the
estimates of fiber orientations in each voxel to form streamlines that traverse
the volume of the white matter \cite{Conturo1999-je, Mori2002-qi}. These methods
have been under increased scrutiny and several lines of investigation have
raised critiques of their validity \cite{Maier-Hein2017-vb, Thomas2014-ki}. On
the other hand, there have been efforts to shore up the inferences made with
these methods \cite{Pestilli2014NatMeth, Takemura2016-sh, Smith2013-nc,
Smith2015-cx, Smith2015-zt, Rheault2018-wk}. Importantly, though discovering
novel fiber bundles requires extraordinary evidence, and delineating the exact
location of the termination of the streamlines in the gray matter, there is
little dispute that tractography can accurately delineate major white matter
bundles that are known to exist within the core of the white matter
\ariel{citation needed}.

@@

In the present work, we present a novel framework for analysis of tractometry
that simultaneously selects the features for analysis, and fits a model to these
features. Rather than a comparative approach, this framework takes a predictive
approach to analysis \cite{Yarkoni2017-bi}. We use a linear modeling approach,
which aims to predict phenotypical variance in a group of subjects, based on a
linear combination of the features estimated with tractometry. Using this
approach, we need to first deal with the large and asymmetric dimensionality of
the data: tractometry data usually has many more features (i.e., number of
measurements per individual) than samples (number of subjects), which makes
inferences from the data about phenotypical differences between individuals
ill-posed. This regime is the target of several statistical learning techniques,
and is often solved by various forms of regularization. For example, Tikhonov
regularization shrinks the solution such that the sum of squared contributions
from the individual features are minimized \cite{Hoerl2000-ij}. Another solution
to the problem is provided by the Lasso algorithm, which instead minimizes the
sum of the absolute values of contributions of each feature
\cite{Tibshirani1996-qs}. This tends to shrink to zero the contributions of many
of the features, providing results that are both accurate and interpretable.
When additional structure is available in the organization of the data,
regularization algorithms can take advantage of this structure. For example, if
the features lend themselves to a natural division into different groups, the
group lasso (GL) can be used to select groups of features, rather than
individual features \cite{Yuan2006-ky}. The Sparse Group Lasso (SGL) elaborates
on this idea by providing control both of group sparsity, as well as overall
sparsity of the solutions \cite{simon2013sgl}. Because the features measured
with tractomery lend themselves to grouping based on the tracts from which each
measurement is taken, GL and SGL provide a useful tool linear model fitting in
problems of this form.

>>>>

Diffusion-weighted Magnetic Resonance Imaging (dMRI) provides a unique view into
the physical properties of the connections that comprise the brain white matter
\cite{wandell2016clarifying}. Among the benefits of the method are the fact that
it is non-invasive, and feasible to conduct in clinical scanners, and in wide
variety of populations: healthy adults, but also clinical populations, as well
as young or elderly subjects. While the measurements are usually conducted with
voxels at the millimeter scale, within the diffusion time usually used, water
molecules within each voxel can probe the structure of white matter at the
micrometer scale, providing important information about the physical structure
of the white matter, such as the distribution of fiber orientations within each
voxel \cite{Frank2002-iz, Frank2001-xf}. With more sophisticated measurements
that include several diffusion-weightings, more information can be extracted
about the composition of different tissue type in the voxel
\cite{Jeurissen2014-ab, Kaden2016-ho}. In addition, phenomenological
descriptions of the diffusion properties, such as its mean diffusivity or
fractional anisotropy can be derived from dMRI measurements
\cite{Pierpaoli1996-vj}. Some of these descriptions are ambiguous in terms of
their interpretation, as variance in these properties can relate to several
different underlying physical causes. For example, fractional anisotropy can
decrease both because of demyelination, but also because of decreased density,
or because of partial volume with other tissue types or between crossing fiber
populations \cite{Beaulieu1996-fn, Beaulieu2002-tl}. Even so, these descriptions
have proven useful in characterizing individual differences or differences
between populations \ariel{citation needed}.

To relate the diffusion in each voxel to the macro-structure of long-range
connections between different brain regions, methods for computational
tract-tracing from diffusion MRI, or tractography, combine the estimates of
fiber orientations in each voxel to form streamlines that traverse the volume of
the white matter \cite{Conturo1999-je, Mori2002-qi}. These methods have been
under increased scrutiny and several lines of investigation have raised
critiques of their validity \cite{Maier-Hein2017-vb, Thomas2014-ki}. On the
other hand, there have been efforts to shore up the inferences made with these
methods \cite{Pestilli2014NatMeth, Takemura2016-sh, Smith2013-nc, Smith2015-cx,
Smith2015-zt, Rheault2018-wk}. Importantly, though discovering novel fiber
bundles requires extraordinary evidence, and delineating the exact location of
the termination of the streamlines in the gray matter, there is little dispute
that tractography can accurately delineate major white matter bundles that are
known to exist within the core of the white matter \cite{Maier-Hein2017-vb}.

@@

In the present work, we present a novel framework for analysis of tractometry
that simultaneously selects the features for analysis, and fits a model to these
features. Rather than a comparative approach, this framework takes a predictive
approach to inference \cite{Breiman2001}. We use a linear modeling approach,
which aims to predict phenotypical variance in a group of subjects, based on a
linear combination of the features estimated with tractometry. Using this
approach, we need to first deal with the large and asymmetric dimensionality of
the data: tractometry data usually has many more features (i.e., number of
measurements per individual) than samples (number of subjects), which makes
inferences from the data about phenotypical differences between individuals
ill-posed. This regime is the target of several statistical learning techniques,
and is often solved by various forms of regularization. For example, Tikhonov
regularization shrinks the solution such that the sum of squared contributions
from the individual features are minimized \cite{Hoerl2000-ij}. Another solution
to the problem is provided by the Lasso algorithm, which instead minimizes the
sum of the absolute values of contributions of each feature
\cite{Tibshirani1996-qs}. This tends to shrink to zero the contributions of many
of the features, providing results that are both accurate and interpretable.
When additional structure is available in the organization of the data,
regularization algorithms can take advantage of this structure. For example, if
the features lend themselves to a natural division into different groups, the
group lasso (GL) can be used to select groups of features, rather than
individual features \cite{Yuan2006-ky}. The Sparse Group Lasso (SGL) elaborates
on this idea by providing control both of group sparsity, as well as overall
sparsity of the solutions \cite{simon2013sgl}. Because the features measured
with tractomery lend themselves to grouping based on the tracts from which each
measurement is taken, GL and SGL provide a useful tool linear model fitting in
problems of this form. Using linear regression as the basis for analysis is
useful for both regression tasks, in which a continuous measure (e.g., age) is
modeled, as well as for classification tasks, where a discrete property is
modeled (e.g., ). Moreover, use of SGL in particular helps derive informative
properties of the human white matter, both in cases where the underlying causes
of variance are sparsely concentrated (e.g., in one particular white matter
tract) and in cases where biological substrates of differences between
individuals are more widespread. This is because SGL is controlled by
meta-parameters that set the degree to  which and that can be set empirically,
based on the data, in a process of nested cross-validation. We demonstrate these
properties of the algorithm in data from previously published studies
\cite{sarica2017corticospinal, yeatman2014lifespan}.
